---
title: "Influenza CITEseq"
author: "Neal Liu"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

CITE-seq single cell data of baseline PBMC samples from 20 healthy individuals (10 high and 10 low responders) 
vaccinated with influenza pandemic H1N1 and seasonal vaccines in 2009.

Citation:
Kotliarov, Y., Sparks, R. et al. Broad immune activation underlies shared set point signatures 
for vaccine responsiveness in healthy individuals and disease activity in patients with lupus. 
Nat. Med. DOI: https://doi.org/10.1038/s41591-020-0769-8 (2020)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries & our own helper functions

```{r}
library(SeuratObject)
library(Seurat)
library(tidyr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(viridis)
library(mclust)
library(edgeR)
library(Rfast)
library(zellkonverter)
library(ADTnorm)
library(dsb)
library(harmony)
library(batchelor)
library(lisi)
library(ggpubr)
library(fastcluster)
#devtools::install('testPackage')
#library(testPackage)
#source('~/Edited_HelperFunctions_scMultiOmics_SuperFastRUVIIIPrPs.R')
source('testPackage/R/Sc_helper_functions.R')
devtools::source_url('https://raw.githubusercontent.com/Neal-Liu1/Multi-omics-integration/main/Helper_functions.R')


```

# Load data from VAST
```{r}
data <- readRDS('/vast/scratch/users/liu.ne/21008967')
data <- SeuratObject::UpdateSeuratObject(object = data)

data2 <- readRDS('/vast/scratch/users/liu.ne/20706642')

data3 <- readRDS('/vast/scratch/users/liu.ne/21347871')

data4 <- readRDS('/vast/scratch/users/liu.ne/20706645')
```

```{r}


```




