---
title: "Influenza CITEseq"
author: "Neal Liu"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

CITE-seq single cell data of baseline PBMC samples from 20 healthy individuals (10 high and 10 low responders) 
vaccinated with influenza pandemic H1N1 and seasonal vaccines in 2009.

Citation:
Kotliarov, Y., Sparks, R. et al. Broad immune activation underlies shared set point signatures 
for vaccine responsiveness in healthy individuals and disease activity in patients with lupus. 
Nat. Med. DOI: https://doi.org/10.1038/s41591-020-0769-8 (2020)


# Load libraries & our own helper functions
```{r}
library(SeuratObject)
library(Seurat)
library(tidyr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(viridis)
library(mclust)
library(edgeR)
library(Rfast)
library(zellkonverter)
library(ADTnorm)
library(dsb)
library(harmony)
library(batchelor)
library(lisi)
library(ggpubr)
library(fastcluster)
library(BiocSingular)
library(parallel)
#devtools::install('testPackage')
#library(testPackage)
#source('~/Edited_HelperFunctions_scMultiOmics_SuperFastRUVIIIPrPs.R')
source('testPackage/R/Sc_helper_functions.R')
devtools::source_url('https://raw.githubusercontent.com/Neal-Liu1/Multi-omics-integration/main/Helper_functions.R')


```

# Load and preprocess data
```{r}
data <- readRDS('/vast/scratch/users/liu.ne/20706645')

# Somehow the old seurat object doesnt line up the cells for RNA and ADT so we need to reorder them
processed_data <- lineup_samples(list(RNA = data@raw.data, ADT = data@assay$CITE@raw.data, Metadata = t(data@meta.data)))

```

# Create Metrics object and run all normalizations
```{r}
RNA_metrics <- BenchmarkMetrics(
  Algorithm = c('Raw_data', 'Seurat_LogNormalize', 'SCTransform','Seurat_CCA', 'Harmony','fastMNN', 'totalVI', 'RUVIII'),
  Raw_data = processed_data[['RNA']],
  Metadata = as.data.frame(t(processed_data[['Metadata']]))
  )

RNA_metrics <- FilterLowlyExpressedGenes(RNA_metrics, batch_variable = 'batch', n_not_detected_batch_to_permit = 0, ngenes_per_batch = 10000)

# RNA_metrics <- NormalizeRNA(RNA_metrics, batch_variable = 'batch')

# RNA_metrics <- ComputeAssessments(RNA_metrics, variables = c('batch', 'K3'))

```


```{r}

test <- RNA_metrics
a <- sample(53000, 20000)
test@Raw_data <- test@Raw_data[,a]
test@Metadata <- test@Metadata[a,]
test@Metadata$log2_libsize <- Rfast::colsums(as.matrix(test@Raw_data))

test <- NormalizeRNA(test)
test <- ComputeAssessments(test, variables = c('batch', 'K3'))

```

```{r, fig.width= 9, fig.height= 3}
variables <- c('batch', 'K3', 'sampleid', 'log2_libsize')
PlotMultipleCorrelations(test, variables = variables, titles = c('Batch', 'Celltype', 'Donor', 'Libsize'))

```

```{r}
p1 <- PlotMultipleLISI(test, 'batch', levels = test@Algorithm, title = 'Batch LISI')
p2 <- PlotMultipleLISI(test, 'K3', levels = test@Algorithm, title = 'Celltype LISI')
gridExtra::grid.arrange(p1,p2, ncol=2)
```
```{r}
p1 <- PlotMultipleSilhouette(test, 'batch', 'boxplot', 'Batch silhouette', 1.2)
p2 <- PlotMultipleSilhouette(test, 'K3', 'boxplot', 'Celltype silhouette', 1.2)
gridExtra::grid.arrange(p1,p2, ncol=2)
```
```{r}
p1 <- PlotARIs(test, 'K3', title = 'Celltype ARI')
p2 <- PlotARIs(test, 'batch', title = 'Batch ARI')
gridExtra::grid.arrange(p1,p2, ncol=2)
```
```{r}
p1 <- PlotHVG_Conservation(test, 'batch', n_top_hvgs = 628)
p2 <- PlotRuntime(test, log=F)
gridExtra::grid.arrange(p1,p2, ncol=2)
```

# fast python umap
```{r}

#reticulate::use_python("/stornext/System/data/apps/python/python-3.11.4/bin/python3.11")
#reticulate::py_install("umap-learn")

env_name <- "/home/users/allstaff/liu.ne/Single-cell-multi-omics-normalization/python_env"
reticulate::virtualenv_create(
  envname = env_name, python = "/stornext/System/data/apps/python/python-3.11.4/bin/python3.11", 
  packages = c("umap-learn", "anndata", "scanpy", "scvi-tools"))

reticulate::use_virtualenv(env_name)

np <- reticulate::import('numpy')
umap <- reticulate::import("umap")
scvi <- reticulate::import('scvi')

```

```{r}

fastumap <- function(matrix, neighbors = 30, min_dist = 0.01, n_components = 2){
  
  # a <- reticulate::r_to_py(list(backend='gpu'))
  py_data <- reticulate::r_to_py(matrix)
  model <- umap$UMAP(
    n_neighbors=neighbors, 
    min_dist= min_dist, 
    n_components= n_components)

  embeddings <- model$fit_transform(py_data)
  embeddings_r <- reticulate::py_to_r(embeddings)
  
  return(embeddings_r)
}



```


```{r}
start <- Sys.time()
fastumap(t(test@Raw_data[1:25,]))
#uwot::umap(as.matrix(t(test@Raw_data[1:25,])))
print(Sys.time()- start)

```



# wilcox test 
```{r}

Rfast::correls(x = test@Adj_data$Seurat_LogNormalize, y = test@Metadata$batch, type = 'spearman')



```



# Backup data
```{r}
# save.image(file= '/vast/scratch/users/liu.ne/Influenza_CITEseq_backup.RData')
# load('/vast/scratch/users/liu.ne/Influenza_CITEseq_backup.RData')

```


```{r}

rownames(dat) <- c("feature1", "feature2", "feature3", "feature4")
colnames(dat) <- paste("Sample", 1:10)

#Calculate UMAP
fast_uamp(dat)


```



