---
title: "Influenza CITEseq"
author: "Neal Liu"
date: "2024-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

CITE-seq single cell data of baseline PBMC samples from 20 healthy individuals (10 high and 10 low responders) 
vaccinated with influenza pandemic H1N1 and seasonal vaccines in 2009.

Citation:
Kotliarov, Y., Sparks, R. et al. Broad immune activation underlies shared set point signatures 
for vaccine responsiveness in healthy individuals and disease activity in patients with lupus. 
Nat. Med. DOI: https://doi.org/10.1038/s41591-020-0769-8 (2020)


# Load libraries & our own helper functions
```{r}
library(SeuratObject)
library(Seurat)
library(tidyr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(viridis)
library(mclust)
library(edgeR)
library(Rfast)
library(zellkonverter)
library(ADTnorm)
library(dsb)
library(harmony)
library(batchelor)
library(lisi)
library(ggpubr)
library(fastcluster)
library(BiocSingular)
library(parallel)
library(rnndescent)
#devtools::install('testPackage')
#library(testPackage)
#source('~/Edited_HelperFunctions_scMultiOmics_SuperFastRUVIIIPrPs.R')
source('testPackage/R/Sc_helper_functions.R')
devtools::source_url('https://raw.githubusercontent.com/Neal-Liu1/Multi-omics-integration/main/Helper_functions.R')


```

# Load and preprocess data
```{r}
# data <- readRDS('/vast/scratch/users/liu.ne/20706645')

# Somehow the old seurat object doesnt line up the cells for RNA and ADT so we need to reorder them
processed_data <- lineup_samples(list(RNA = data@raw.data, ADT = data@assay$CITE@raw.data, Metadata = t(data@meta.data)))

```

# Create Metrics object and run all normalizations
```{r}
RNA_metrics <- BenchmarkMetrics(
  Algorithm = c('Raw_data', 'Seurat_LogNormalize', 'SCTransform','Seurat_CCA', 
                'Harmony','fastMNN', 'totalVI', 'RUVIII'),
  Raw_data = processed_data[['RNA']],
  Metadata = as.data.frame(t(processed_data[['Metadata']]))
  )

RNA_metrics <- FilterLowlyExpressedGenes(
  RNA_metrics, batch_variable = 'batch', 
  n_not_detected_batch_to_permit = 0, ngenes_per_batch = 10000)

# RNA_metrics <- NormalizeRNA(RNA_metrics, batch_variable = 'batch')
# RNA_metrics <- ComputeAssessments(RNA_metrics, variables = c('batch', 'K3'))

```


```{r}

test <- RNA_metrics
a <- sample(53000, 20000)
test@Raw_data <- test@Raw_data[,a]
test@Metadata <- test@Metadata[a,]
test@Metadata$log2_libsize <- log2(colSums(test@Raw_data))
```

```{r}
test <- NormalizeRNA(test)
```


# RUVIII
# Unsupervised PRPC
```{r}
test_seurat <- Seurat::CreateSeuratObject(
  processed_data$RNA[,a], assay = 'RNA', 
  meta.data = data.frame(t(processed_data$Metadata[,a])),
  key = 'rna_')

test_seurat@assays$ADT <- Seurat::CreateAssayObject(processed_data$ADT[,a], key = 'adt_')
test_seurat$log2_rna_libsize <- log2(colSums(test_seurat@assays$RNA@layers$counts))

test_seurat <- FindCorrectedMultimodalNeighbours(
  test_seurat, assays = c('RNA', 'ADT'), 
  batch_variable = 'batch', 
  normalization_methods = c('LogNormalize','CLR'),
  npcs = c(30,18))

```

```{r}
# Loading matrixExtra gives weird the no slot i in Graph error
test_seurat <- Seurat::FindClusters(
  test_seurat, algorithm = 2, graph.name = 'wsnn', 
  cluster.name = 'harmony-wnn', resolution = 1)

mclust::adjustedRandIndex(test_seurat$`harmony-wnn`, test_seurat$K3)
```

# Make PRPC
```{r}

rna_counts <- processed_data$RNA[,a]
rownames(rna_counts) <- rownames(processed_data$RNA)
rna_counts <- rna_counts[test@Raw_data@Dimnames[[1]],]

# batch prpc
batch_prpc <- createPrPc_default(
  rna_counts, test_seurat$`harmony-wnn`, 
  test_seurat$batch, sampling = 2, colname_suffix = 'batch')

# donor prpc
donor_prpc <- createPrPc_default(
  rna_counts, test_seurat$`harmony-wnn`,
  test_seurat$sampleid, sampling = 3, colname_suffix = 'donor')

# calculate libsize 
test_seurat$log2_rna_libsize <- log2(colSums(test_seurat@assays$RNA@layers$counts))

ls_prpc_batch1 <- createPrPc_default(
  rna_counts[, which(test_seurat$batch == 1)], 
  test_seurat$`harmony-wnn`[which(test_seurat$batch == 1)], 
  test_seurat$log2_rna_libsize[which(test_seurat$batch == 1)], 
  sampling = 4, colname_suffix = 'ls_batch1', continuous_bins = 3)

ls_prpc_batch2 <- createPrPc_default(
  rna_counts[, which(test_seurat$batch == 2)], 
  test_seurat$`harmony-wnn`[which(test_seurat$batch == 2)], 
  test_seurat$log2_rna_libsize[which(test_seurat$batch == 2)], 
  sampling = 4, colname_suffix = 'ls_batch2', continuous_bins = 3)

ls_prpc_all <- createPrPc_default(
  rna_counts, test_seurat$`harmony-wnn`, 
  test_seurat$log2_rna_libsize, 
  sampling = 6, colname_suffix = 'ls', 
  continuous_bins = 2, separate_bins_by_biology = T)

prpc <- cbind(batch_prpc, donor_prpc, ls_prpc_batch1, ls_prpc_batch2)

```

# Find ncg
```{r}
test_singlecellexperiment <- SingleCellExperiment(
  assays= list(RNA = as.matrix(rna_counts)), colData= test_seurat@meta.data)

ncg <- RUVIIIPRPS::findNcgSupervised(
  test_singlecellexperiment, assay.name = 'RNA', bio.variables = 'harmony-wnn', 
  uv.variables = c('log2_rna_libsize'), assess.se.obj = F, save.se.obj = F)

```

```{r}
test <- RunRUVIII(
  test, prpc, ncg, k = 4, celltype = 'harmony-wnn', 
  M = ruv::replicate.matrix(colnames(prpc)), pcs = 30)

test <- RunRUVIII(
  test, donor_prpc, ncg, k = 2, celltype = 'harmony-wnn', 
  M = ruv::replicate.matrix(colnames(donor_prpc)), pcs = 30)

test <- RunRUVIII(
  test, cbind(ls_prpc_batch1, ls_prpc_batch2), ncg, k = 2, celltype = 'harmony-wnn', 
  M = ruv::replicate.matrix(colnames(cbind(ls_prpc_batch1, ls_prpc_batch2))), pcs = 30)

test <- RunRUVIII(
  test, ls_prpc_all, ncg, k = 1, celltype = 'harmony-wnn', 
  M = ruv::replicate.matrix(colnames(ls_prpc_all)), pcs = 30)

```

```{r, fig.width= 9, fig.height= 3}
variables <- c('batch', 'K3', 'sampleid', 'log2_libsize')
PlotMultipleCorrelations(test, variables = variables, titles = c('Batch', 'Celltype', 'Donor', 'Libsize'))

```

```{r}
plot_PCA(test@Adj_data$RUVIII, 'ruvIII', test@Metadata$K3, 'celltypes', is_pca_obj = F, pcs_to_compute = 2)
plot_PCA(test@Adj_data$RUVIII, 'ruvIII', test@Metadata$log2_libsize, 'libsize', is_pca_obj = F, pcs_to_compute = 2, is_continuous = T)
plot_PCA(test@Adj_data$RUVIII, 'ruvIII', test@Metadata$sampleid, 'donor', is_pca_obj = F, pcs_to_compute = 2, is_continuous = F)
plot_PCA(test@Adj_data$RUVIII, 'ruvIII', test@Metadata$batch, 'batch', is_pca_obj = F, pcs_to_compute = 2, is_continuous = F)
```

```{r}
test <- ComputeAssessments(test, variables = c('batch', 'K3'))

```

```{r}
p1 <- PlotMultipleLISI(test, 'batch', levels = test@Algorithm, title = 'Batch LISI')
p2 <- PlotMultipleLISI(test, 'K3', levels = test@Algorithm, title = 'Celltype LISI')
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
p1 <- PlotMultipleSilhouette(test, 'batch', 'boxplot', 'Batch silhouette', 1.2)
p2 <- PlotMultipleSilhouette(test, 'K3', 'boxplot', 'Celltype silhouette', 1.2)
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
p1 <- PlotARIs(test, 'K3', title = 'Celltype ARI')
p2 <- PlotARIs(test, 'batch', title = 'Batch ARI')
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
p1 <- PlotHVG_Conservation(test, 'batch', n_top_hvgs = 800)
p2 <- PlotRuntime(test, log=F)
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
test <- ComputeUMAP(test)

```

```{r}
plot_UMAP(test@UMAPs$, test@Metadata$batch, run_umap = F, alpha = 0.35)
plot_UMAP(test@UMAPs$Seurat_CCA, test@Metadata$K3, run_umap = F, alpha = 0.35)

```

```{r}


```

# wilcox test 
```{r}
hist(wilcoxon_test(as.matrix(test@Raw_data), test@Metadata$batch, n.cores = 90)$Adj.pvalue)
hist(wilcoxon_test(as.matrix(test@Adj_data$Seurat_LogNormalize), test@Metadata$batch, n.cores = 90)$Adj.pvalue)
#hist(wilcoxon_test(as.matrix(test@Adj_data$Seurat_CCA), test@Metadata$batch, n.cores = 90)$Adj.pvalue)
hist(wilcoxon_test(as.matrix(test@Adj_data$SCTransform), test@Metadata$batch, n.cores = 90)$Adj.pvalue)
hist(wilcoxon_test(as.matrix(test@Adj_data$RUVIII), test@Metadata$batch, n.cores = 90)$Adj.pvalue)

```

```{r}
ADT_metrics <- BenchmarkMetrics(
  Raw_data = processed_data[['ADT']][!grepl('sotype_', processed_data[['ADT']]@Dimnames[[1]]),],
  Metadata = as.data.frame(t(processed_data[['Metadata']]))
  )
ADT_metrics@Algorithm <- c('Raw_counts', 'DSB', 'CLR', 'Harmony', 'ADTnorm', 'totalVI', 'RUVIII')

ADT_metrics <- NormalizeADT(ADT_metrics, params = list(), batch_variable = 'batch')

```

# RUV normalization
```{r}




```

# Plot correlations
```{r, fig.width= 9, fig.height= 3}
variables <- c('batch', 'K3', 'sampleid', 'log2_ADT_libsize')
PlotMultipleCorrelations(ADT_metrics, variables = variables, titles = c('Batch', 'Celltype', 'Donor', 'Libsize'))

```



```{r}
ADT_metrics <- ComputeAssessments(ADT_metrics, c('batch', 'K3'))

```

```{r}
p1 <- PlotMultipleLISI(ADT_metrics, 'batch', levels = ADT_metrics@Algorithm, title = 'Batch LISI')
p2 <- PlotMultipleLISI(ADT_metrics, 'K3', levels = ADT_metrics@Algorithm, title = 'Celltype LISI')
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
p1 <- PlotMultipleSilhouette(ADT_metrics, 'batch', 'boxplot', 'Batch silhouette', 1.2)
p2 <- PlotMultipleSilhouette(ADT_metrics, 'K3', 'boxplot', 'Celltype silhouette', 1.2)
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
p1 <- PlotARIs(ADT_metrics, 'K3', title = 'Celltype ARI')
p2 <- PlotARIs(ADT_metrics, 'batch', title = 'Batch ARI')
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r}
PlotRuntime(ADT_metrics, log=F)
```




# Backup data
```{r}
# save.image(file= '/vast/scratch/users/liu.ne/Influenza_CITEseq_backup.RData')
load('/vast/scratch/users/liu.ne/Influenza_CITEseq_backup.RData')

```



