---
title: "Influenza CITEseq"
author: "Neal Liu"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import libraries
```{r}
library(Seurat)
library(tidyr)
library(ggplot2)
library(viridis)
library(cowplot)
library(igraph)
library(tidyverse)

source("~/scMultiOmics-normalization/Sc_helper_functions.R")
devtools::source_url('https://raw.githubusercontent.com/Neal-Liu1/Multi-omics-integration/main/Helper_functions.R')

```


# Import data
Citation:
Kotliarov, Y., Sparks, R. et al. Broad immune activation underlies shared set point signatures 
for vaccine responsiveness in healthy individuals and disease activity in patients with lupus. 
Nat. Med. DOI: https://doi.org/10.1038/s41591-020-0769-8 (2020)

CITE-seq single cell data of baseline PBMC samples from 20 healthy individuals (10 high and 10 low responders) 
vaccinated with influenza pandemic H1N1 and seasonal vaccines in 2009.

```{r}
# Read data
# wget https://nih.figshare.com/articles/dataset/CITE-seq_protein-mRNA_single_cell_data_from_high_and_low_vaccine_responders_to_reproduce_Figs_4-6_and_associated_Extended_Data_Figs_/11349761/1?file=20706645

data <- readRDS('/vast/scratch/users/liu.ne/20706645')

# Somehow the old seurat object doesnt line up the cells for RNA and ADT so we need to reorder them
raw_data_matrices <- lineup_samples(list(RNA = data@raw.data, ADT = data@assay$CITE@raw.data, Metadata = t(data@meta.data)))

#raw_data_matrices <- readRDS('/vast/scratch/users/liu.ne/influenza_rawdata.rds')

```

# Load and prepare datasets
```{r}
metadata <- raw_data_matrices$Metadata %>% t() %>% as.data.frame()
rna_counts <- raw_data_matrices$RNA %>% as('dgCMatrix')
adt_counts <- raw_data_matrices$ADT %>% as('dgCMatrix')
metadata$log2_rna_libsize <- colSums(rna_counts) %>% log2()
metadata$log2_adt_libsize <- colSums(adt_counts) %>% log2()

rna_metrics <- BenchmarkMetrics(
  Algorithm = c('Raw_counts', 'Seurat_LogNormalize', 'SCTransform','Seurat_rPCA',
                'Harmony_LogNormalize','fastMNN', 'totalVI', 
                'fastRUVIII_supervised', 'fastRUVIII_unsupervised'), 
  Raw_data = rna_counts, 
  Metadata = metadata
)

adt_metrics <- BenchmarkMetrics(
  Algorithm = c('Raw_counts', 'ADTnorm', 'CLR', 'DSB', 'Harmony_CLR', 
                'totalVI', 'fastRUVIII_supervised', 'fastRUVIII_unsupervised'),
  Raw_data = adt_counts[!grepl('sotype_', adt_counts@Dimnames[[1]]),], 
  Metadata = metadata)

rna_metrics <- FilterLowlyExpressedGenes(
  rna_metrics, 'batch', ngenes_per_batch = 10000, 
  n_not_detected_batch_to_permit = 0)

seurat_obj <- Seurat::CreateSeuratObject(counts = rna_counts, meta.data = metadata)
seurat_obj@assays$ADT <- Seurat::CreateAssayObject(counts = adt_counts, key = 'adt_')

# totalVI python conversion shenanigans
test <- SingleCellExperiment::SingleCellExperiment(assays = list(ADT = adt_metrics@Raw_data),
                             colData = metadata)
test2 <- SingleCellExperiment::SingleCellExperiment(assays = list(RNA = rna_metrics@Raw_data),
                             colData = metadata)
# Write H5
zellkonverter::writeH5AD(test, file = '/vast/scratch/users/liu.ne/ADT_influenza.h5ad')
zellkonverter::writeH5AD(test2, file = '/vast/scratch/users/liu.ne/RNA_influenza.h5ad')
rm(test)
rm(test2)

```

# Run the other normalizations
```{r}
rna_metrics <- NormalizeRNA(rna_metrics)
adt_metrics <- NormalizeADT(adt_metrics, num_pcs = 15)

```

# get totalVI data
# only run this after totalVI python code has been run
```{r}
totalVI_adt <- zellkonverter::readH5AD('/vast/scratch/users/liu.ne/totalVI_adt_influenza.h5ad')
totalVI_rna <- zellkonverter::readH5AD('/vast/scratch/users/liu.ne/totalVI_rna_influenza.h5ad')
rna_metrics <- add_totalVI(rna_metrics, totalVI_rna, assay_name = 'rna_normalized', pca = T)
adt_metrics <- add_totalVI(adt_metrics, totalVI_adt, assay_name = 'adt_normalized', pcs = 15)

```

# Unsupervised prpc for fastRUVIII
```{r}
seurat_obj <- FindCorrectedMultimodalNeighbours(
  seurat_obj, assays = c('RNA', 'ADT'), uv_variables = c('batch', 'sampleid'), 
  npcs = c(30, 15), normalization_methods = c('LogNormalize', 'CLR'), 
  graph.name = 'harmony_wsnn')

seurat_obj <- FindClusters(
  seurat_obj, graph.name = 'harmony_wsnn', resolution = 0.5, 
  cluster.name = 'harmony_wnn_clusters', algorithm = 3)

rna_metrics@Metadata <- seurat_obj@meta.data
adt_metrics@Metadata <- seurat_obj@meta.data

mclust::adjustedRandIndex(seurat_obj$K3, seurat_obj$harmony_wnn_clusters)

```

# Make prpcs
```{r}
rna_donor_prpc_supervised <- createPrPc_default(
  rna_metrics@Raw_data, rna_metrics@Metadata$K3, 
  uv_vector = rna_metrics@Metadata$sampleid, colname_suffix = 'donor')

rna_donor_prpc_unsupervised <- createPrPc_default(
  rna_metrics@Raw_data, rna_metrics@Metadata$harmony_wnn_clusters, 
  uv_vector = rna_metrics@Metadata$sampleid, colname_suffix = 'donor')

adt_donor_prpc_supervised <- createPrPc_default(
  adt_metrics@Raw_data, adt_metrics@Metadata$K3, 
  uv_vector = adt_metrics@Metadata$sampleid, colname_suffix = 'donor')

adt_donor_prpc_unsupervised <- createPrPc_default(
  adt_metrics@Raw_data, adt_metrics@Metadata$harmony_wnn_clusters, 
  uv_vector = adt_metrics@Metadata$sampleid, colname_suffix = 'donor')

rna_ls_prpc_supervised <- createPrPc_default(
  rna_metrics@Raw_data, rna_metrics@Metadata$K3, 
  uv_vector = rna_metrics@Metadata$log2_rna_libsize, colname_suffix = 'ls')

rna_ls_prpc_unsupervised <- createPrPc_default(
  rna_metrics@Raw_data, rna_metrics@Metadata$harmony_wnn_clusters, 
  uv_vector = rna_metrics@Metadata$log2_rna_libsize, colname_suffix = 'ls')

adt_ls_prpc_supervised <- createPrPc_default(
  adt_metrics@Raw_data, adt_metrics@Metadata$K3, 
  uv_vector = adt_metrics@Metadata$log2_adt_libsize, colname_suffix = 'ls')

adt_ls_prpc_unsupervised <- createPrPc_default(
  adt_metrics@Raw_data, adt_metrics@Metadata$harmony_wnn_clusters, 
  uv_vector = adt_metrics@Metadata$log2_adt_libsize, colname_suffix = 'ls')
```


# Find ncgs
```{r}
# donor_ncg_supervised <- FindNCG(
#   rna_metrics, unwanted_variables = 'sampleid', 
#   bio_variables = 'K3', no.ncg = 500)
# donor_ncg_unsupervised <- FindNCG(
#   rna_metrics, unwanted_variables = 'sampleid', 
#   bio_variables = 'harmony_wnn_clusters', no.ncg = 500)
# 
# ls_ncg_supervised <- FindNCG(
#   rna_metrics, unwanted_variables = 'log2_rna_libsize', 
#   bio_variables = 'K3', no.ncg = 500)
# ls_ncg_unsupervised <- FindNCG(
#   rna_metrics, unwanted_variables = 'log2_rna_libsize', 
#   bio_variables = 'harmony_wnn_clusters', no.ncg = 500)

all_ncg_supervised <- FindNCG(
  rna_metrics, unwanted_variables = c('log2_rna_libsize'), 
  bio_variables = c('sampleid','K3'), no.ncg = 600)
all_ncg_unsupervised <- FindNCG(
  rna_metrics, unwanted_variables = c('log2_rna_libsize'), 
  bio_variables = c('sampleid', 'harmony_wnn_clusters'), no.ncg = 600)

```

# Run the fastRUVIIIs
```{r}
# Only adjusting donor
"
rna_metrics <- RunRUVIII(
  rna_metrics, prpc = rna_donor_prpc_supervised, 
  ncgs = rownames(rna_metrics@Raw_data) %in% donor_ncg_supervised, 
  name = 'fastRUVIII_supervised_donor', 
  k = 1)


rna_metrics <- RunRUVIII(
  rna_metrics, prpc = rna_donor_prpc_unsupervised, 
  ncgs = rownames(rna_metrics@Raw_data) %in% donor_ncg_unsupervised, 
  name = 'fastRUVIII_unsupervised_donor', 
  k = 3)
"
# Only adjusting libsize
rna_metrics <- RunRUVIII(
  rna_metrics, prpc = rna_ls_prpc_supervised, 
  ncgs = rownames(rna_metrics@Raw_data) %in% ls_ncg_supervised, 
  name = 'fastRUVIII_supervised', 
  k = 1)


rna_metrics <- RunRUVIII(
  rna_metrics, prpc = rna_ls_prpc_unsupervised, 
  ncgs = rownames(rna_metrics@Raw_data) %in% ls_ncg_unsupervised, 
  name = 'fastRUVIII_unsupervised', 
  k = 1)
"

# Adjusting both
rna_metrics <- RunRUVIII(
  rna_metrics, prpc = cbind(rna_ls_prpc_supervised, rna_donor_prpc_supervised), 
  ncgs = rownames(rna_metrics@Raw_data) %in% all_ncg_supervised, 
  name = 'fastRUVIII_supervised', 
  k = 1)

rna_metrics <- RunRUVIII(
  rna_metrics, prpc = cbind(rna_ls_prpc_unsupervised, rna_donor_prpc_unsupervised), 
  ncgs = rownames(rna_metrics@Raw_data) %in% all_ncg_unsupervised, 
  name = 'fastRUVIII_unsupervised', 
  k = 1)

rna_metrics <- RunRUVIII(
  rna_metrics, prpc = cbind(rna_ls_prpc_supervised, rna_donor_prpc_supervised), 
  ncgs = rownames(rna_metrics@Raw_data) %in% all_ncg_unsupervised, 
  name = 'fastRUVIII_unsupervised_all', 
  k = 3)

rna_metrics@Algorithm <- c(
  'Raw_counts', 'Seurat_LogNormalize', 'SCTransform','Seurat_rPCA',
  'Harmony_LogNormalize','fastMNN', 'totalVI', 
  'fastRUVIII_supervised_donor', 'fastRUVIII_supervised_libsize', 
  'fastRUVIII_supervised_all')
"


```


```{r}
"
adt_metrics <- RunRUVIII(
  adt_metrics, prpc = adt_donor_prpc_supervised, 
  ncgs = rep(T, nrow(adt_metrics@Raw_data)), 
  name = 'fastRUVIII_supervised_donor', 
  k = 1)

adt_metrics <- RunRUVIII(
  adt_metrics, prpc = adt_ls_prpc_supervised, 
  ncgs = rep(T, nrow(adt_metrics@Raw_data)), 
  name = 'fastRUVIII_supervised_libsize', 
  k = 1)

"
adt_metrics <- RunRUVIII(
  adt_metrics, prpc = adt_ls_prpc_supervised, 
  ncgs = rep(T, nrow(adt_metrics@Raw_data)), 
  name = 'fastRUVIII_supervised', 
  k = 5)

adt_metrics <- RunRUVIII(
  adt_metrics, prpc = adt_ls_prpc_unsupervised, 
  ncgs = rep(T, nrow(adt_metrics@Raw_data)), 
  name = 'fastRUVIII_unsupervised', 
  k = 5)

"
adt_metrics@Algorithm <- c(
  'Raw_counts', 'ADTnorm', 'CLR', 'DSB', 'Harmony_CLR', 
  'totalVI', 'fastRUVIII_supervised_donor', 
  'fastRUVIII_supervised_libsize', 
  'fastRUVIII_supervised_all')
"



```


```{r, fig.width= 12, fig.height= 3}
p1 <- PlotMultipleCorrelations(rna_metrics, c('batch', 'K3', 'sampleid', 'log2_rna_libsize'), titles = c('Batch','Celltype', 'Donor', 'Libsize'))
p1
```


```{r, fig.width= 12, fig.height= 3}
p1 <- PlotMultipleCorrelations(adt_metrics, c('batch', 'K3', 'sampleid', 'log2_adt_libsize'), titles = c('Batch','Celltype', 'Donor', 'Libsize'))
p1
```

```{r}
time1 <- Sys.time()
adt_metrics <- ComputeMultipleLISI(adt_metrics, variables = c('sampleid', 'K3'))
paste0(
  "ADT LISI completed in ",
  as.numeric(difftime(Sys.time(), time1, units = "mins")),
  " minutes"
)
rna_metrics <- ComputeMultipleLISI(rna_metrics, variables = c('sampleid', 'K3'))
paste0(
  "RNA LISI completed in ",
  as.numeric(difftime(Sys.time(), time1, units = "mins")),
  " minutes"
)
adt_metrics <- ComputeARIs(
  adt_metrics, 'K3')
rna_metrics <- ComputeARIs(
  rna_metrics, 'K3')

paste0(
  "Assessments completed in ",
  as.numeric(difftime(Sys.time(), time1, units = "mins")),
  " minutes"
)

```

```{r, fig.width= 12}
gradient_palette <- colorRampPalette(c("#e7ebf7", "#90add9"))
colors <- gradient_palette(length(rna_metrics@Algorithm))

p1 <- PlotMultipleLISI(rna_metrics, "K3", levels = rna_metrics@Algorithm, title = "RNA celltyle LISI")+
  scale_fill_manual(values = colors)
p2 <- PlotMultipleLISI(adt_metrics, "K3", levels = rna_metrics@Algorithm, title = "ADT celltype LISI")+
  scale_fill_manual(values = colors)
p3 <- PlotMultipleLISI(rna_metrics, "sampleid", levels = rna_metrics@Algorithm, title = "RNA donor LISI")+
  scale_fill_manual(values = colors)
p4 <- PlotMultipleLISI(adt_metrics, "sampleid", levels = rna_metrics@Algorithm, title = "ADT donor LISI")+
  scale_fill_manual(values = colors)

gridExtra::grid.arrange(p1, p2, p3 , p4, ncol = 4)

```

# Plot Clustering metrics
```{r, fig.width= 12}
library(ggplot2)
library(gridExtra)

# RNA plot
df_rna <- data.frame(
  Method = rna_metrics@ARI$K3$results_df$method,
  AMI_celltype = unlist(rna_metrics@ARI$K3$results_df$AMI),
  ARI_celltype = unlist(rna_metrics@ARI$K3$results_df$ARI),
  knn = rna_metrics@ARI$K3$results_df$k,
  res = rna_metrics@ARI$K3$results_df$gamma
)

p4 <- ggplot(df_rna, aes(x = AMI_celltype, y = ARI_celltype)) +
  geom_point(aes(color = Method), size = 2.5, alpha = 0.4) +
  
  # highlight fastRUVIII_supervised
  geom_point(
    data = subset(df_rna, Method == "fastRUVIII_supervised"),
    color = "black",
    size = 4.5, alpha = 0.9
  ) + 
  geom_point(
    data = subset(df_rna, Method == "fastRUVIII_supervised"),
    aes(color = Method),
    size = 4, alpha = 0.6
  ) +
  geom_point(
    data = subset(df_rna, Method == "fastRUVIII_unsupervised"),
    color = "black",
    size = 4.5, alpha = 0.9
  ) + 
  # highlight fastRUVIII_unsupervised
  geom_point(
    data = subset(df_rna, Method == "fastRUVIII_unsupervised"),
    aes(color = Method),
    size = 4, alpha = 0.6
  ) +
  
  theme_minimal(base_size = 14) +
  labs(
    y = "ARI",
    x = "AMI",
    title = "AMI & ARI of RNA Methods on celltypes"
  ) +
  theme(
    aspect.ratio = 1,
    panel.border = element_rect(colour = "grey55", fill = NA, size = 1)
  )

# ADT plot
df_adt <- data.frame(
  Method = adt_metrics@ARI$K3$results_df$method,
  AMI_celltype = unlist(adt_metrics@ARI$K3$results_df$AMI),
  ARI_celltype = unlist(adt_metrics@ARI$K3$results_df$ARI),
  knn = adt_metrics@ARI$K3$results_df$k,
  res = adt_metrics@ARI$K3$results_df$gamma
)

p5 <- ggplot(df_adt, aes(x = AMI_celltype, y = ARI_celltype)) +
  geom_point(aes(color = Method), size = 2.5, alpha = 0.4) +
  
  # highlight fastRUVIII_supervised
  geom_point(
    data = subset(df_adt, Method == "fastRUVIII_supervised"),
    color = "black",
    size = 4.5, alpha = 0.9
  ) + 
  geom_point(
    data = subset(df_adt, Method == "fastRUVIII_supervised"),
    aes(color = Method),
    size = 4, alpha = 0.6
  ) +
  geom_point(
    data = subset(df_adt, Method == "fastRUVIII_unsupervised"),
    color = "black",
    size = 4.5, alpha = 0.9
  ) + 
  # highlight fastRUVIII_unsupervised
  geom_point(
    data = subset(df_adt, Method == "fastRUVIII_unsupervised"),
    aes(color = Method),
    size = 4, alpha = 0.6
  ) +
  
  theme_minimal(base_size = 14) +
  labs(
    y = "ARI",
    x = "AMI",
    title = "AMI & ARI of ADT Methods on celltypes"
  ) +
  theme(
    aspect.ratio = 1,
    panel.border = element_rect(colour = "grey55", fill = NA, size = 1)
  )

gridExtra::grid.arrange(p4, p5, ncol = 2)

```

```{r}
PlotRuntime(obj = rna_metrics, title = 'RNA runtimes') +
  coord_cartesian(ylim = c(0,5))
PlotRuntime(obj = adt_metrics, title = 'ADT runtimes') +
  coord_cartesian(ylim = c(0,5))

```


```{r}
adt_metrics <- ComputeUMAP(adt_metrics, method = 'uwot')
rna_metrics <- ComputeUMAP(rna_metrics, method = 'uwot')

```


```{r,fig.width= 11, fig.height= 4}
# > names(rna_metrics@UMAPs)
# [1] "Raw_counts"              "Seurat_LogNormalize"     "fastMNN"                 "Harmony_LogNormalize"   
# [5] "SCTransform"             "Seurat_rPCA"             "totalVI"                 "fastRUVIII_supervised"  
# [9] "fastRUVIII_unsupervised"

method = "fastRUVIII_unsupervised"
method2 = "fastRUVIII_supervised" 

p1 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = paste0(method,' libsize'), run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 11.8)

p2 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$K3, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

p3 <- plot_UMAP(
  rna_metrics@UMAPs[[method2]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = paste0(method2,' libsize'), run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 11.8)

p4 <- plot_UMAP(
  rna_metrics@UMAPs[[method2]], metadata_vector = rna_metrics@Metadata$K3, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1, p2,p3,p4, ncol = 4)

```


```{r, fig.width= 11, fig.height= 4}

# > names(adt_metrics@UMAPs)
# [1] "Raw_counts"              "CLR"                     "Harmony_CLR"             "DSB"                    
# [5] "ADTnorm"                 "totalVI"                 "fastRUVIII_supervised"   "fastRUVIII_unsupervised"

method = "fastRUVIII_supervised"
method2 = "fastRUVIII_unsupervised"

p1 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = paste0(method,' libsize'), run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 11.8)

p2 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$K3, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

p3 <- plot_UMAP(
  adt_metrics@UMAPs[[method2]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = paste0(method2,' libsize'), run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 11.8)

p4 <- plot_UMAP(
  adt_metrics@UMAPs[[method2]], metadata_vector = rna_metrics@Metadata$K3, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1, p2,p3,p4, ncol = 4)

```

```{r}


```



```{r}
# save.image(file= '/vast/scratch/users/liu.ne/paper_panels_3_backup.RData')
# load('/vast/scratch/users/liu.ne/paper_panels_3_backup.RData')

```










