---
title: "Influenza CITEseq"
author: "Neal Liu"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import libraries
```{r}
#Sys.setenv(RETICULATE_PYTHON = "~/.venvs/leiden/bin/python")
library(Seurat)
library(tidyverse)
library(ggplot2)
library(gridExtra)
devtools::source_url('https://raw.githubusercontent.com/Neal-Liu1/Multi-omics-integration/main/Helper_functions.R')
source("~/scMultiOmics-normalization/Sc_helper_functions.R")

```

# Import Stuart et al's bone marrow CITE-seq
# Citation: 
Stuart T, Butler A, Hoffman P, Hafemeister C, Papalexi E, Mauck WM 3rd, Hao Y, Stoeckius M, Smibert P, Satija R. 
Comprehensive Integration of Single-Cell Data. Cell. 2019 Jun 13;177(7):1888-1902.e21. 
doi: 10.1016/j.cell.2019.05.031.
```{r}
# SeuratData::InstallData("bmcite", destdir = '/vast/scratch/users/liu.ne')
merged_data <- readRDS('/vast/scratch/users/liu.ne/combined_seurat.rds')
bm_data <- merged_data

```








# DATA SETUP
```{r}

triana_2021 <- readRDS(file = '/vast/scratch/users/liu.ne/41038073')

DefaultAssay(triana_2021) <- 'RNA'
triana_2021 <- NormalizeData(triana_2021) %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA(reduction.name='rna_pca')
triana_2021 <- Seurat::RunUMAP(triana_2021, reduction = 'rna_pca', dims = 1:30, reduction.name = 'rna_umap')

DefaultAssay(triana_2021) <- 'AB'
# VariableFeatures(triana_2021) <- rownames(triana_2021[["AB"]])
triana_2021 <- RunPCA(triana_2021, reduction.name = 'adt_pca')
triana_2021 <- RunUMAP(triana_2021, reduction = 'adt_pca', reduction.name = 'adt_umap', dims = 1:30)


```

## Filtering genes from both studies
```{r}
bm_genes_to_keep <- edgeR::filterByExpr(as.matrix(bm_data@assays$RNA@layers$counts), group= bm_data$celltype.l2, min.count = 1, min.total.count = 20)

triana_genes_to_keep <- edgeR::filterByExpr(as.matrix(triana_2021@assays$RNA@counts), group= triana_2021$ct, min.count = 3.5, min.total.count = 20)

common_genes <- intersect(rownames(bm_data@assays$RNA)[bm_genes_to_keep] ,rownames(triana_2021@assays$RNA@counts)[triana_genes_to_keep])

```

## Do SCTransform first on both datasets before label transfer
```{r}

bm_data2 <- CreateSeuratObject(counts = bm_data1@assays$RNA@layers$counts[rownames(bm_data@assays$RNA) %in% common_genes,])
rownames(bm_data2) <- rownames(bm_data@assays$RNA)[rownames(bm_data@assays$RNA) %in% common_genes]
bm_data2 <- Seurat::SCTransform(bm_data2)
bm_data2 <- FindVariableFeatures(bm_data2)

triana_2021_2 <- CreateSeuratObject(counts = triana_2021@assays$RNA@counts[rownames(triana_2021@assays$RNA@counts) %in% common_genes,], 
                                    meta.data = triana_2021@meta.data)

triana_2021_2 <- Seurat::SCTransform(triana_2021_2)

```

## Label transfer
```{r}

# For L1 celltypes
anchors <- FindTransferAnchors(reference = bm_data2, 
                               normalization.method = 'SCT',
                               query = triana_2021_2, 
                               dims = 1:30)

predicted_celltypes <- TransferData(anchorset = anchors, refdata = bm_data$celltype.l1, dims = 1:30)
triana_2021_2@meta.data$celltype.l1_with_transfer <- predicted_celltypes$predicted.id


# For l2 celltypes
predicted_celltypes2 <- TransferData(anchorset = anchors, refdata = bm_data$celltype.l2, dims = 1:30)
triana_2021_2@meta.data$celltype.l2_with_transfer <- predicted_celltypes2$predicted.id


# Plot max prediction scores to see how confident the predictions are
hist(predicted_celltypes$prediction.score.max, main = 'Max predict. scores for l1 celltypes')
hist(predicted_celltypes2$prediction.score.max, main = 'Max predict. scores for l2 celltypes')

```

## Merge the two studies into one Seurat
```{r}

# Deal with metadata shenanigans
bm_data2$study <- 'Stuart_2019'
triana_2021_2@assays$ADT@key <- 'adt_'
triana_2021_2$study <- 'Triana_2021'
bm_data2$supplied_celltype_l2 <- as.vector(bm_data$celltype.l2)
triana_2021_2$supplied_celltype_l2 <- as.vector(triana_2021$ct)
bm_data2$celltype.l1_with_transfer <- as.vector(bm_data$celltype.l1)
bm_data2$celltype.l2_with_transfer <- as.vector(bm_data$celltype.l2)

# merge 
merged_data <- merge(bm_data2, triana_2021_2)

```













# Run the other normalization methods for RNA and ADT
```{r}
bm_data$log2_rna_libsize <- log2(colSums(bm_data@assays$RNA@layers$counts))
bm_data$log2_adt_libsize <- log2(colSums(bm_data@assays$ADT@layers$counts))

"
genes_to_keep <- Seurat::FindVariableFeatures(
  as.matrix(bm_data@assays$RNA@layers$counts), nfeatures = 8000, 
  selection.method = 'vst')[['variable']]
"
rna_metrics <- BenchmarkMetrics(
  Algorithm = c('Raw_counts', 'Seurat_LogNormalize', 'SCTransform','Seurat_rPCA',
                'Harmony_LogNormalize','fastMNN', 'totalVI', 
                'fastRUVIII_supervised', 'fastRUVIII_unsupervised'), 
  Raw_data = bm_data@assays$RNA@layers$counts, 
  Metadata = bm_data@meta.data)
rownames(rna_metrics@Raw_data) <- dimnames(bm_data@assays$RNA)[[1]]
colnames(rna_metrics@Raw_data) <- dimnames(bm_data@assays$RNA)[[2]]

adt_metrics <- BenchmarkMetrics(
  Algorithm = c('Raw_counts', 'ADTnorm', 'CLR', 'DSB', 'Harmony_CLR', 
                'totalVI', 'fastRUVIII_supervised', 'fastRUVIII_unsupervised'),
  Raw_data = bm_data@assays$ADT@layers$counts, 
  Metadata = bm_data@meta.data)
rownames(adt_metrics@Raw_data) <- dimnames(bm_data@assays$ADT)[[1]]
colnames(adt_metrics@Raw_data) <- dimnames(bm_data@assays$ADT)[[2]]

# totalVI python conversion shenanigans
test <- SingleCellExperiment::SingleCellExperiment(assays = list(ADT = adt_metrics@Raw_data),
                             colData = bm_data@meta.data)
test2 <- SingleCellExperiment::SingleCellExperiment(assays = list(RNA = rna_metrics@Raw_data),
                             colData = bm_data@meta.data)
# Write H5
zellkonverter::writeH5AD(test, file = '/vast/scratch/users/liu.ne/ADT_merged.h5ad')
zellkonverter::writeH5AD(test2, file = '/vast/scratch/users/liu.ne/RNA_merged.h5ad')
rm(test)
rm(test2)

# Run all other methods for RNA and ADT
rna_metrics <- NormalizeRNA(
  rna_metrics, batch_variable = 'study', num_pcs = 30)

adt_metrics <- NormalizeADT(
  adt_metrics, batch_variable = 'study', num_pcs = 13)

```


# Setting up the universal colour palette for consistency across plots
```{r}

# Manual color scheme
method_cols <- c(
  # Shared across modalities
  Raw_counts              = "#7A7A7A",  # neutral grey
  totalVI                 = "#CC79A7",  # magenta (stands out, consistent)
  fastRUVIII_supervised   = "#D55E00",  # vermillion
  fastRUVIII_unsupervised = "#E69F00",  # orange

  # RNA-specific
  Seurat_LogNormalize     = "#0072B2",  # blue
  SCTransform             = "#56B4E9",  # light blue
  Seurat_rPCA             = "#009E73",  # green
  Harmony_LogNormalize    = "#1B9E10",  # teal-green (distinct from Seurat_rPCA)
  fastMNN                 = "#F0E442",  # yellow (high visibility)

  # ADT-specific
  ADTnorm                 = "#332288",  # deep indigo
  CLR                     = "#117730",  # deep green (different family than Seurat_rPCA)
  DSB                     = "#AA4499",  # purple
  Harmony_CLR             = "#44AA99"   # aqua/teal (ADT harmony variant)
)

# Auto palette color scheme
method_cols_hue <- scales::hue_pal()(length(method_cols))
names(method_cols_hue) <- names(method_cols)



```



# Unsupervised prpc for fastRUVIII
```{r}
bm_data <- FindCorrectedMultimodalNeighbours(
  bm_data, assays = c('RNA', 'ADT'), uv_variables = c('study'), 
  npcs = c(30, 18), normalization_methods = c('LogNormalize', 'CLR'), 
  graph.name = 'harmony_wsnn')

bm_data <- FindClusters(
  bm_data, graph.name = 'harmony_wsnn', resolution = 0.8, 
  cluster.name = 'harmony_wnn_clusters', algorithm = 3)

rna_metrics@Metadata <- bm_data@meta.data
adt_metrics@Metadata <- bm_data@meta.data

mclust::adjustedRandIndex(bm_data$celltype.l2_with_transfer, bm_data$harmony_wnn_clusters)

```

# Get prpc and ncgs
```{r}
supervised_prpc_rna <- CreatePRPC(
  rna_metrics@Raw_data, bio_vars = 'celltype.l2_with_transfer', uv_vars = c('log2_rna_libsize','study'), 
  group_by_vars = c('study', NA), separate_bins_by_biology = c(T, NA), continuous_bins = 3, 
  metadata = rna_metrics@Metadata)

unsupervised_prpc_rna <- CreatePRPC(
  rna_metrics@Raw_data, bio_vars = 'harmony_wnn_clusters', uv_vars = c('log2_rna_libsize','study'), 
  group_by_vars = c('study', NA), separate_bins_by_biology = c(T, NA), continuous_bins = 3, 
  metadata = bm_data@meta.data)

supervised_prpc_adt <- CreatePRPC(
  adt_metrics@Raw_data, bio_vars = 'celltype.l2_with_transfer', uv_vars = c('log2_adt_libsize','study'), 
  group_by_vars = c('study', NA), separate_bins_by_biology = c(T, NA), continuous_bins = 3, 
  metadata = adt_metrics@Metadata)

unsupervised_prpc_adt <- CreatePRPC(
  adt_metrics@Raw_data, bio_vars = 'harmony_wnn_clusters', uv_vars = c('log2_adt_libsize','study'), 
  group_by_vars = c('study', NA), separate_bins_by_biology = c(T, NA), continuous_bins = 3, 
  metadata = adt_metrics@Metadata)

# NCGs
supervised_ncg <- FindNCG(
  rna_metrics, unwanted_variables = c('log2_rna_libsize', 'study'), 
  bio_variables = 'celltype.l2_with_transfer', no.ncg = 800, apply_log = T, 
  sample_fraction = 0.15)

unsupervised_ncg <- FindNCG(
  rna_metrics, unwanted_variables = c('log2_rna_libsize', 'study'), 
  bio_variables = 'harmony_wnn_clusters', no.ncg = 800, apply_log = T, 
  sample_fraction = 0.15)


```

```{r}
plot_linear_correlation(
  list(log2_sparse(rna_metrics@Raw_data[supervised_ncg,])), c('ncg_pca'), 
  rna_metrics@Metadata$log2_rna_libsize, 'libsize', num_pcs = 5, is.pca.obj = F)

plot_vector_correlation(
  list(log2_sparse(rna_metrics@Raw_data[supervised_ncg,])), c('ncg_pca'), 
  rna_metrics@Metadata$study, 'batch', num_pcs = 5, is.pca.obj = F)

plot_vector_correlation(
  list(log2_sparse(rna_metrics@Raw_data[supervised_ncg,])), c('ncg_pca'), 
  rna_metrics@Metadata$celltype.l2_with_transfer, 'celltype', num_pcs = 5, is.pca.obj = F)

plot_linear_correlation(
  list(log2_sparse(rna_metrics@Raw_data[unsupervised_ncg,])), c('ncg_pca'), 
  rna_metrics@Metadata$log2_rna_libsize, 'libsize', num_pcs = 5, is.pca.obj = F)

plot_vector_correlation(
  list(log2_sparse(rna_metrics@Raw_data[unsupervised_ncg,])), c('ncg_pca'), 
  rna_metrics@Metadata$study, 'batch', num_pcs = 5, is.pca.obj = F)

plot_vector_correlation(
  list(log2_sparse(rna_metrics@Raw_data[unsupervised_ncg,])), c('ncg_pca'), 
  rna_metrics@Metadata$celltype.l2_with_transfer, 'celltype', num_pcs = 5, is.pca.obj = F)


```

# Add totalVI data after we have the totalVI results
```{r}
totalVI_adt <- zellkonverter::readH5AD('/vast/scratch/users/liu.ne/totalVI_adt_merged.h5ad')
totalVI_rna <- zellkonverter::readH5AD('/vast/scratch/users/liu.ne/totalVI_rna_merged.h5ad')
rna_metrics <- add_totalVI(rna_metrics, totalVI_rna, assay_name = 'rna_normalized', pca = T)
adt_metrics <- add_totalVI(adt_metrics, totalVI_adt, assay_name = 'adt_normalized', pcs = 13)

# totalVI_rna_counts <- rna_metrics@Adj_data$totalVI
# rna_metrics@Adj_data$totalVI <- log2(totalVI_rna_counts+1e-8)
rna_metrics@PCs$totalVI <- BiocSingular::runSVD(t(rna_metrics@Adj_data$totalVI), 30, center = T, BSPARAM = BiocSingular::bsparam())$u
# 
# totalVI_adt_counts <- adt_metrics@Adj_data$totalVI
# adt_metrics@Adj_data$totalVI <- log2(totalVI_adt_counts+1e-8)
adt_metrics@PCs$totalVI <- BiocSingular::runSVD(t(adt_metrics@Adj_data$totalVI), 13, center = T, BSPARAM = BiocSingular::bsparam())$u
```

# Run fastRUVIII
```{r}
rna_metrics <- RunRUVIII(
  rna_metrics, prpc = supervised_prpc_rna, ncgs = rownames(rna_metrics@Raw_data) %in% supervised_ncg, 
  k = 20, name = 'fastRUVIII_supervised')

rna_metrics <- RunRUVIII(
  rna_metrics, prpc = unsupervised_prpc_rna, ncgs = rownames(rna_metrics@Raw_data) %in% unsupervised_ncg, 
  k = 20, name = 'fastRUVIII_unsupervised')

adt_metrics <- RunRUVIII(
  adt_metrics, prpc = supervised_prpc_adt, ncgs = rep(T, nrow(adt_metrics@Raw_data)), 
  k = 2, name = 'fastRUVIII_supervised', pcs = 13)

adt_metrics <- RunRUVIII(
  adt_metrics, prpc = unsupervised_prpc_adt, ncgs = rep(T, nrow(adt_metrics@Raw_data)), 
  k = 2, name = 'fastRUVIII_unsupervised', pcs = 13)

```


# Try scMerge2
```{r}

# if(!"scMerge2" %in% rna_metrics@Algorithm)
#   {rna_metrics@Algorithm <- c(rna_metrics@Algorithm, "scMerge2")}
# 
# scmerge_res <- scMerge::scMerge2(
#   log2(rna_metrics@Raw_data+1), 
#   cellTypes = rna_metrics@Metadata$celltype.l2_with_transfer, 
#   batch = rna_metrics@Metadata$study,
#   ruvK =  20)
# 
# rna_metrics@Adj_data$scMerge2 <- scmerge_res$newY %>% as("dgCMatrix")
# rna_metrics@PCs$scMerge2 <- run_PCA(scmerge_res$newY, pcs = 30)$u

# rna_metrics@Adj_data$scMerge2 <- NULL
# rna_metrics@PCs$scMerge2 <- NULL
# rna_metrics@Algorithm <- rna_metrics@Algorithm[1:9]


```



# Plot metrics
```{r, fig.height= 6, fig.width= 10, dev='svg'}
p1 <- PlotMultipleCorrelations(
  rna_metrics,
  c("study", "celltype.l2_with_transfer", "log2_rna_libsize"),
  titles = c("Batch", "Celltype", "Libsize"),
  num_pcs = 10,
  palette = method_cols_hue
)

p2 <- PlotMultipleCorrelations(
  adt_metrics,
  c("study", "celltype.l2_with_transfer", "log2_adt_libsize"),
  titles = c("Batch", "Celltype", "Libsize"),
  num_pcs = 10,
  palette = method_cols_hue
)

gridExtra::grid.arrange(p1,p2, nrow = 2)

```



# Compute assessments (This might take a while, maybe one or two hours)
```{r}
adt_metrics <- ComputeAssessments(adt_metrics, c('celltype.l2_with_transfer', 'study'))
rna_metrics <- ComputeAssessments(rna_metrics, c('celltype.l2_with_transfer', 'study'))

adt_metrics <- ComputeARIs(
  adt_metrics, 'celltype.l2_with_transfer')

adt_metrics <- ComputeARIs(
  adt_metrics, 'study')

rna_metrics <- ComputeARIs(
  rna_metrics, 'celltype.l2_with_transfer')

rna_metrics <- ComputeARIs(
  rna_metrics, 'study')

```


```{r, fig.width= 6, fig.height= 4}
gradient_palette <- colorRampPalette(c("#e7ebf7", "#90add9"))
colors <- gradient_palette(length(rna_metrics@Algorithm))

# p3 <- PlotARIs(adt_metrics, c('celltype.l2_with_transfer'), title = 'ADT Celltype ARI') + scale_fill_manual(values = colors)
# p4 <- PlotMultipleSilhouette(adt_metrics, variable = 'celltype.l2_with_transfer', title = 'Silhouettes per celltype', plot_type = 'boxplot')
# p4 <- PlotARIs(adt_metrics, c('study'), title = 'Batch ARI') + scale_fill_manual(values = colors)
p5 <- PlotMultipleLISI(adt_metrics, variable = c('celltype.l2_with_transfer'), title = 'ADT Celltype LISI', levels = adt_metrics@Algorithm) + 
  scale_fill_manual(values = colors)
p6 <- PlotMultipleLISI(adt_metrics, variable = c('study'), title = 'ADT Batch LISI', levels = adt_metrics@Algorithm) +
  scale_fill_manual(values = colors)
gridExtra::grid.arrange(p5, p6, ncol = 2)
```


```{r, fig.width= 6, fig.height= 4}
# p3 <- PlotARIs(rna_metrics, c('celltype.l2_with_transfer'), title = 'RNA Celltype ARI') + 
#   scale_fill_manual(values = colors)
# p4 <- PlotARIs(rna_metrics, c('study'), title = 'Batch ARI') +
#   scale_fill_manual(values = colors)

p5 <- PlotMultipleLISI(rna_metrics, variable = c('celltype.l2_with_transfer'), title = 'RNA Celltype LISI', levels = adt_metrics@Algorithm) +
  scale_fill_manual(values = colors)
p6 <- PlotMultipleLISI(rna_metrics, variable = c('study'), title = 'RNA Batch LISI', levels = adt_metrics@Algorithm) +
  scale_fill_manual(values = colors)
gridExtra::grid.arrange(p5,p6, ncol = 2)
```


# Plot Clustering metrics
```{r,fig.width= 12}
df_ct_rna <- data.frame(
  Method = rna_metrics@ARI$celltype.l2_with_transfer$results_df$method,
  AMI_celltype = unlist(rna_metrics@ARI$celltype.l2_with_transfer$results_df$AMI),
  ARI_celltype = unlist(rna_metrics@ARI$celltype.l2_with_transfer$results_df$ARI),
  knn = rna_metrics@ARI$celltype.l2_with_transfer$results_df$k,
  res = rna_metrics@ARI$celltype.l2_with_transfer$results_df$gamma
)

df_batch_rna <- data.frame(
  Method = rna_metrics@ARI$study$results_df$method,
  AMI_celltype = unlist(rna_metrics@ARI$study$results_df$AMI),
  ARI_celltype = unlist(rna_metrics@ARI$study$results_df$ARI),
  knn = rna_metrics@ARI$study$results_df$k,
  res = rna_metrics@ARI$study$results_df$gamma
)

p4_rna <- plot_ami_ari(df_ct_rna, "AMI & ARI of RNA Methods on celltypes", palette = method_cols_hue)
p5_rna <- plot_ami_ari(df_batch_rna, "AMI & ARI of RNA Methods on Batch", palette = method_cols_hue)

gridExtra::grid.arrange(p4_rna, p5_rna, ncol = 2)


```


```{r, fig.width= 12}

df_ct_adt <- data.frame(
  Method = adt_metrics@ARI$celltype.l2_with_transfer$results_df$method,
  AMI_celltype = unlist(adt_metrics@ARI$celltype.l2_with_transfer$results_df$AMI),
  ARI_celltype = unlist(adt_metrics@ARI$celltype.l2_with_transfer$results_df$ARI),
  knn = adt_metrics@ARI$celltype.l2_with_transfer$results_df$k,
  res = adt_metrics@ARI$celltype.l2_with_transfer$results_df$gamma
)

df_batch_adt <- data.frame(
  Method = adt_metrics@ARI$study$results_df$method,
  AMI_celltype = unlist(adt_metrics@ARI$study$results_df$AMI),
  ARI_celltype = unlist(adt_metrics@ARI$study$results_df$ARI),
  knn = adt_metrics@ARI$study$results_df$k,
  res = adt_metrics@ARI$study$results_df$gamma
)

p4_adt <- plot_ami_ari(df_ct_adt, "AMI & ARI of ADT Methods on celltypes", palette = method_cols_hue)
p5_adt <- plot_ami_ari(df_batch_adt, "AMI & ARI of ADT Methods on Batch", palette = method_cols_hue)

gridExtra::grid.arrange(p4_adt, p5_adt, ncol = 2)




```



```{r}
PlotRuntime(obj = rna_metrics, title = 'RNA runtimes') +
  coord_cartesian(ylim = c(0,5))
PlotRuntime(obj = adt_metrics, title = 'ADT runtimes') +
  coord_cartesian(ylim = c(0,5))
PlotRuntime(obj = rna_metrics, title = 'RNA runtimes')
PlotRuntime(obj = adt_metrics, title = 'RNA runtimes')

```

# UMAPs 
```{r}
adt_metrics <- ComputeUMAP(adt_metrics, method = 'uwot')
rna_metrics <- ComputeUMAP(rna_metrics, method = 'uwot')

```

```{r, fig.width= 10, fig.height= 4}
method = 'Raw_counts'

p1 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = 'Raw counts: libsize', run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 12.7)
p2 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$study, 
  title = 'Batch', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$celltype.l2_with_transfer, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3, ncol = 3)

```



```{r, fig.width= 10, fig.height= 4}
# These need to be required in the function
library(viridis)
library(cowplot)

method = 'scMerge2'

p1 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = 'scMerge2: libsize', run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 12.7)
p2 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$study, 
  title = 'Batch', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$celltype.l2_with_transfer, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3, ncol = 3)

```


```{r, fig.width= 10, fig.height= 4}

method = 'fastRUVIII_unsupervised'

p1 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = 'fastRUVIII_unsupervised: libsize', run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 12.7)
p2 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$study, 
  title = 'Batch', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$celltype.l2_with_transfer, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3, ncol = 3)

```

```{r, fig.width= 10, fig.height= 4}

method = 'totalVI'

p1 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$log2_rna_libsize, 
  title = 'totalVI: libsize', run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 12.7)
p2 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$study, 
  title = 'Batch', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  rna_metrics@UMAPs[[method]], metadata_vector = rna_metrics@Metadata$celltype.l2_with_transfer, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3, ncol = 3)

```


```{r, fig.width= 10, fig.height= 4}
method = 'fastRUVIII_unsupervised'

p1 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = adt_metrics@Metadata$log2_adt_libsize, 
  title = 'fastRUVIII_unsupervised ADT: libsize', run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 13.5)
p2 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = adt_metrics@Metadata$study, 
  title = 'Batch', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = adt_metrics@Metadata$celltype.l2_with_transfer, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3, ncol = 3)

```

```{r, fig.width= 10, fig.height= 4}
method = 'totalVI'

p1 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = adt_metrics@Metadata$log2_adt_libsize, 
  title = 'totalVI ADT: libsize', run_umap = F, label_is_continuous = T, alpha = 0.4, continuous_var_upper_lim = 13.5)
p2 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = adt_metrics@Metadata$study, 
  title = 'Batch', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  adt_metrics@UMAPs[[method]], metadata_vector = adt_metrics@Metadata$celltype.l2_with_transfer, 
  title = 'Celltype', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3, ncol = 3)

```



```{r}
unsupervised_clustering_umaps <- BenchmarkMetrics(
  Raw_data = rna_metrics@Raw_data,
  Metadata = rna_metrics@Metadata,
  PCs = list(bm_data@reductions$pca_RNA@cell.embeddings, 
             bm_data@reductions$pca_ADT@cell.embeddings, 
             bm_data@reductions$harmony_RNA@cell.embeddings, 
             bm_data@reductions$harmony_ADT@cell.embeddings))

unsupervised_clustering_umaps <- ComputeUMAP(
  unsupervised_clustering_umaps, method = 'uwot')

```



```{r, fig.width= 9, fig.height= 11}
p1 <- plot_UMAP(
  unsupervised_clustering_umaps@UMAPs[[1]], metadata_vector = unsupervised_clustering_umaps@Metadata$s, 
  title = 'Log RNA', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p2 <- plot_UMAP(
  unsupervised_clustering_umaps@UMAPs[[2]], metadata_vector = unsupervised_clustering_umaps@Metadata$celltype.l2_with_transfer, 
  title = 'CLR ADT', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p3 <- plot_UMAP(
  unsupervised_clustering_umaps@UMAPs[[3]], metadata_vector = unsupervised_clustering_umaps@Metadata$celltype.l2_with_transfer, 
  title = 'Batch corrected RNA', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p4 <- plot_UMAP(
  unsupervised_clustering_umaps@UMAPs[[4]], metadata_vector = unsupervised_clustering_umaps@Metadata$celltype.l2_with_transfer, 
  title = 'Batch corrected ADT', run_umap = F, label_is_continuous = F, alpha = 0.4, show_density = F)
p5 <- plot_UMAP(
  bm_data@reductions$wnn_normalized_umap@cell.embeddings, 
  metadata_vector = unsupervised_clustering_umaps@Metadata$study, 
  title = 'WNN integration', run_umap = F, label_is_continuous = F, 
  alpha = 0.4, show_density = F)

gridExtra::grid.arrange(p1,p2,p3,p4,p5, ncol = 2, nrow = 3)


```


```{r}
ruv_obj <- Sparse_RUV_III(
  log2_sparse(t(rna_metrics@Raw_data)),
  Yrep =  t(supervised_prpc_rna), 
  M = ruv::replicate.matrix(colnames(supervised_prpc_rna)), 
  ctl = rownames(rna_metrics@Raw_data) %in% supervised_ncg, 
  k = 20, 
  return.info = T)

bad_ncgs <- FindNCG(rna_metrics, unwanted_variables = 'celltype.l2_with_transfer', bio_variables = c('study', 'log2_rna_libsize'))

ruv_obj <- Sparse_RUV_III(
  log2_sparse(t(rna_metrics@Raw_data)),
  Yrep =  t(supervised_prpc_rna), 
  M = ruv::replicate.matrix(colnames(supervised_prpc_rna)), 
  ctl = rownames(rna_metrics@Raw_data) %in% bad_ncgs, 
  k = 20, 
  return.info = T)


```


```{r, fig.width= 5, fig.height= 3}
#plot(x= rna_metrics@Metadata$log2_rna_libsize, y = ruv_obj$W[,1])
#plot(x= rna_metrics@Metadata$study, y = ruv_obj$W[,1])

dim = 8
sample <- sample(1:nrow(ruv_obj$W), 10000)
  
ggplot(
  mapping = aes(
    y = ruv_obj$W[sample,dim],
    x = 1:length(sample), 
    color = rna_metrics@Metadata[['log2_rna_libsize']][sample])) +
  geom_point(alpha = 0.5, size = 0.9) +
  scale_color_viridis(option = "viridis") +
  theme_minimal() +
  labs(x = "Cells", y = "W score", color = 'Libsize (log2)') +
  theme(panel.border=element_rect(colour = "grey60", fill=NA, size=1.5),
        aspect.ratio = 1/2,
        legend.position = 'none')

ggplot(
  mapping = aes(
    y = ruv_obj$W[sample,dim],
    x = 1:length(sample), 
    color = rna_metrics@Metadata[['celltype.l2_with_transfer']][sample])) +
  geom_point(alpha = 0.5, size = 0.9) +
  theme_minimal() +
  labs(x = "Cells", y = "W score", color = 'Celltype') +
  theme(panel.border=element_rect(colour = "grey60", fill=NA, size=1.5), 
        aspect.ratio = 1/2, 
        legend.position = 'none') # Optional: remove y-axis ticks for cells

ggplot(
  mapping = aes(
    y = ruv_obj$W[sample,dim],
    x = 1:length(sample), 
    color = rna_metrics@Metadata[['study']][sample])) +
  geom_point(alpha = 0.5, size = 0.9) +
  theme_minimal() +
  labs(x = "Cells", y = "W score", color = 'Batch') +
  theme(panel.border=element_rect(colour = "grey60", fill=NA, size=1.5), 
        aspect.ratio = 1/2, 
        legend.position = 'none') # Optional: remove y-axis ticks for cells


```


```{r}
W <- ruv_obj$W[1:7,1:20]
rownames(W) <- c('Libsize', 'Batch', '??', '??', '??', '??', 'Biology?')
colnames(W) <- dimnames(rna_metrics@Raw_data)[[1]][1:20]

# Create heatmap without dendrograms and with labels on the left
pheatmap::pheatmap(
  W,
  cluster_rows = FALSE,      # Remove row dendrogram
  cluster_cols = FALSE,      # Remove column dendrogram
  cellwidth = 14,
  cellheight = 14,
  show_rownames = TRUE,  # Ensures row names are displayed on the left
  show_colnames = TRUE
)

```



```{r}
plot_ridgeplot(t(adt_metrics@Adj_data$ADTnorm), rownames(adt_metrics@Raw_data))
plot_ridgeplot(t(adt_metrics@Adj_data$CLR), rownames(adt_metrics@Raw_data))
plot_ridgeplot(t(adt_metrics@Adj_data$DSB), rownames(adt_metrics@Raw_data))
plot_ridgeplot(log2(t(adt_metrics@Adj_data$totalVI)+1), rownames(adt_metrics@Raw_data))
plot_ridgeplot(t(adt_metrics@Adj_data$fastRUVIII_unsupervised), rownames(adt_metrics@Raw_data))

```


```{r, fig.height= 5, fig.width= 8}
protein = 'CD4'
p1 <- plot_ridgeplot(t(adt_metrics@Adj_data$ADTnorm[protein,]), adt_metrics@Metadata$celltype.l2_with_transfer, title = 'ADTnorm')
plot_ridgeplot(t(adt_metrics@Adj_data$CLR[protein,]), adt_metrics@Metadata$celltype.l2_with_transfer)
plot_ridgeplot(t(adt_metrics@Adj_data$DSB[protein,]), adt_metrics@Metadata$celltype.l2_with_transfer)
plot_ridgeplot(t(adt_metrics@Adj_data$totalVI[protein,]), adt_metrics@Metadata$celltype.l2_with_transfer)
p2 <- plot_ridgeplot(t(adt_metrics@Adj_data$fastRUVIII_unsupervised[protein,]), adt_metrics@Metadata$celltype.l2_with_transfer, title = 'fastRUVIII')
rm(protein)

gridExtra::grid.arrange(p1, p2, ncol = 2)

```


```{r}
protein = 'CD11a'
plot_ridgeplot(adt_metrics@Adj_data$ADTnorm[protein,], adt_metrics@Metadata$study)
plot_ridgeplot(adt_metrics@Adj_data$CLR[protein,], adt_metrics@Metadata$study)
plot_ridgeplot(adt_metrics@Adj_data$DSB[protein,], adt_metrics@Metadata$study)
plot_ridgeplot(log2(1+adt_metrics@Adj_data$totalVI[protein,]), adt_metrics@Metadata$study)
plot_ridgeplot(adt_metrics@Adj_data$fastRUVIII_unsupervised[protein,], adt_metrics@Metadata$study)
rm(protein)

```


```{r}
library(plotly)
ruv_obj$u <- ruv_obj$W[,1:3]
colnames(ruv_obj$u) <- c('PC1', 'PC2', 'PC3')
plot_3D_PCA(
  ruv_obj, is_PCA_obj = T, title = 'W matrix', 
  label_vector = rna_metrics@Metadata$celltype.l2_with_transfer)

```


```{r}

df <- rna_metrics@UMAPs$totalVI
df$metadata <- rna_metrics$

ggplot(df, aes(x = UMAP1, y = UMAP2, color = metadata)) +
      geom_point(size = 0.07, alpha = alpha) +
      ggtitle(title) +
      theme_minimal() +
      theme(axis.line = element_line(colour = "grey50", linewidth = 0.9),
            panel.border = element_blank(),  #element_rect(colour = "grey90", fill=NA, size=0.7),
            panel.grid.major = element_blank(),  #element_line(color = "grey96"),
            panel.grid.minor = element_blank(),
            aspect.ratio = aspect_ratio,
            legend.position = "none")+
      geom_text(data = centroids, aes(label = metadata), size = 3, color = "black", hjust = 0.5, vjust = 0.5)


```


```{r}
# save.image(file= '/vast/scratch/users/liu.ne/paper_panels_2_backup.RData')
# load('/vast/scratch/users/liu.ne/paper_panels_2_backup.RData')

# saveRDS(adt_metrics, file = '/vast/scratch/users/liu.ne/adt_metrics_merged.RData')
# saveRDS(list(rna_metrics = rna_metrics), file = '/vast/scratch/users/liu.ne/adt_metrics_merged.RData')

```







